---
title: "bird_joins"
author: "Adelaide Robinson"
date: '2022-08-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(stringr)
library(lubridate)
```

Read in the data sets and store the data frames as bird_observations, sites, surveys, and taxalist (it should be clear from the raw file names which is which) Create a subset of bird_observations called birds_subset that only contains observations for birds with species id "BHCO" and "RWBL", and from sites with site ID "LI-W" and "NU-C"

```{r}
bird_observations <- read_csv(here("data", "knb-lter-cap.256.9", "52_pp52_birds_1.csv"))
sites <- read_csv(here("data", "knb-lter-cap.256.9", "52_pp52_sites_1.csv"))
surveys <- read_csv(here("data", "knb-lter-cap.256.9", "52_pp52_surveys_1.csv"))
taxalist <- read_csv(here("data", "knb-lter-cap.256.9", "52_pp52_taxalist_1.csv"))

birds_subset <- bird_observations |> filter(species_id %in% c("BHCO", "RWBL"), 
                                            site_id %in% c("LI-W", "NU-C") )
```

Use left join(s) to update birds_subset so that it also includes sites and taxalist information. For each join, include an explicit argument saying which variable you are joining by (even if it will just assume the correct one for you). Store the updated data frame as birds_left. Make sure to look at the output - is what it contains consistent with what you expected it to contain?

```{r}

birds_lefta <- left_join(x = birds_subset, y = sites, by = "site_id")

birds_left <-  left_join(x = birds_lefta, y = taxalist, by = "species_id")
```

First, answer: what do you expect a full_join() between birds_subset and sites to contain?

Write code to full_join the birds_subset and sites data into a new object called birds_full. Explicitly include the variable you're joining by. Look at the output. Is it what you expected? I expect a full join to include all observations from birds_subset and sites, giving nas where they do not overlap

```{r}
birds_full <- full_join(birds_subset, sites, by = "site_id")

```

First, answer: what do you expect an inner_join() between birds_subset and taxalist to contain? I expect this to contain all observations that birds_subset and taxalist have, where the species id is included in both datasets

Write code to inner_join() the birds_subset and taxalist, called birds_inner. Include an argument for what variable you'll be joining by.

What would you get if instead of inner_join() you'd used left_join() for this example? Try it out & check. Why does that make sense for this scenario? It what case would you expect the outcome to differ from an inner_join()?

This returns the same dataset, that is because all of the species are included both on the birds_subset and in the taxalist dataframe. The outcome would differ for inner join, if a species_id was included in either dataset that was not included in the other.

```{r}

birds_inner <- inner_join(x = birds_subset, y = taxalist, by = "species_id")

birds_inner2 <- inner_join(x = birds_subset, y = taxalist, by = "species_id")
```

```{r}
birds <- bird_observations |> rename(birds_obs_notes = notes)

new_sub <- birds |> left_join(sites, by = "site_id") |> 
  left_join(taxalist) |>
  left_join(surveys) |> 
  select(survey_date, common_name, park_name, bird_count) |> 
  mutate(survey_month = month(survey_date)) |> 
  mutate(survey_month = factor(survey_month)) |> 
  relocate(survey_month, .after = survey_date) |> 
  group_by(park_name, survey_month) |> 
  summarise(total_birds = sum(bird_count, na.rm = TRUE)) |> 
  filter(park_name %in% c("Lindo","Orme", "Palomino", "Sonrisa"))

```
